---
title: "Cluster-wise regression"
author: "Sherry Liang"
date: "3/3/2018"
output: html_document
---
#### Import data and select features

I selected 8 variables for further analysis: Duration,InstallmentRatePercentage,ResidenceDuration,Age,NumberExistingCredits,NumberPeopleMaintenance,Telephone,ForeignWorker

```{r,warning=FALSE}
load("~/.RData") 
library(rpart)
library(flexmix)
source("~/Documents/clustreg.txt")
source("~/Documents/clustreg.predict.txt")

train0<-read.csv(file="~/Documents/Analytics/Data_Mining/Assignment4/Training700.csv")
holdout0<-read.csv(file="~/Documents/Analytics/Data_Mining/Assignment4/Holdout300.csv")
# select numeric variables and set Amount as the 1st column
train <- train0[c(3,2,4:10)]
holdout <- holdout0[c(3,2,4:10)]
```

#### Build Model

```{r}
clustreg.1<-clustreg(train,1,1,830,1)
clustreg.2<-clustreg(train,2,25,830,12)
clustreg.3<-clustreg(train,3,25,830,12)

plot(c(1,2,3),c(clustreg.1$rsq.best,clustreg.2$rsq.best,clustreg.3$rsq.best),ylim=c(0,1),type="l",col=4,main="VAF Plot for German Credit Data: Cluster-wise Regression", ylab="Variance Accounted For",xlab="Number of Clusters")  
```

> From the plot above we can see that the Variance Accounted For by the model increases as the number of cluster increases.

```{r}
clustreg.2$results[[1]]
```
> There are 4 variables statistically significant in the model with 2 clusters: Duration, InstallmentRatePercentage, Age and Telephone.

```{r}
clustreg.3$results[[1]]
```

> There are also 4 variables statistically significant in the model with 3 clusters: Duration, InstallmentRatePercentage, Telephone and ForeignWorker. Age is no longer statistically significant but ForeignWorker is.

```{r}
cbind(Cluster2.R2=clustreg.2$results[[1]]$r.squared,Cluster2.AdjR2=clustreg.2$results[[1]]$adj.r.squared,Cluster3.R2=clustreg.3$results[[1]]$r.squared,Cluster3.AdjR2=clustreg.3$results[[1]]$adj.r.squared)
```
> The model with 3 clusters has the best performance of R2 and adjusted-R2. Therefore I tend to choose the clustering regression model with 3 clusters. But we still need to perform holdout validation.

```{r}
ho.2=clustreg.predict(clustreg.2, holdout)
ho.2$rsq
table((ho.2$cluster))
round(prop.table(table((ho.2$cluster))),3)
```

```{r}
ho.3=clustreg.predict(clustreg.3, holdout)
ho.3$rsq
table((ho.3$cluster))
round(prop.table(table((ho.3$cluster))),3)
```

> The R2 of the holdout dataset of 3 clusters are still better than the 2 clusters. The distribution of observations of the 2 clusters model is 78.3% and 21.7%. For model with 3 clusters, each cluster account for 56.3%, 11.0% and 32.7%. Although neither is equally distributed, the 3 clusters model is still better. Therefore I chose the model with 3 clusters.
